Modelling BCG vaccination in the UK

### 2019-08-05 rmds/01-introduction.Rmd
updated tb epi figures

* **`idmodelr`**: An R package that contains a library  of infectious disease models as well as modelling utilities. It provides tooling that includes: example SEI/SEIR/SHLIR/SHLITR model code, a model solving wrapper; a model summary function; a scenario analysis function). Used by the explore infectious disease model dashboard (http://seabbs.co.uk/shiny/exploreidmodels/) for all functionality. Available from CRAN using `install.packages("idmodelr")` or install the development version from GitHub with `devtools::install_github("seabbs/idmodelr")`. Link: https://www.samabbott.co.uk/idmodelr/

>>>>

* **`idmodelr`**: An R package that contains a library  of infectious disease models as well as modelling utilities. It provides tooling that includes: example SEI/SEIR/SHLIR/SHLITR model code, a model solving wrapper; a model summary function; and a scenario analysis function. Used by the explore infectious disease model dashboard (http://seabbs.co.uk/shiny/exploreidmodels/) for all functionality. Available from CRAN using `install.packages("idmodelr")` or install the development version from GitHub with `devtools::install_github("seabbs/idmodelr")`. Link: https://www.samabbott.co.uk/idmodelr/

### 2019-08-05 rmds/04-tb-epi-england.Rmd
direct effects hannah comments
updated tb epi figures

Although the characteristics of tuberculosis (TB) in England have been reported elsewhere,[@PHE2017; @PHE2018] and key risk factors such as non-UK birth status have been identified,@French2007 little attention has been given to the role of BCG vaccination. In particular, there is little information available regarding the demographics of vaccinated versus unvaccinated cases and the impact of BCG vaccination on TB outcomes in England has not been explored. There has also only been limited reporting of the age distribution, and trends over time, in incidence rates stratified by UK birth status.

In this chapter I explore the epidemiology of TB in England using routine data-sets, with a particular focus on the impact of missing data, the mechanisms underlying that missing data, seasonal trends, the role of age, UK birth status and BCG status. I have also estimated incidence rates, stratified by UK birth status and age, which I then used to identify trends in TB incidence over time. Finally I report TB outcomes in England using case rates, again stratified by BCG status and UK birth status. These data will then be used throughout this thesis to explore the impact of BCG vaccination on TB outcomes (Chapter \@ref(beneficial-bcg-out)), to estimate the direct impact of the 2005 change in BCG vaccination policy (Chapter \@ref(direct-eff)), to parameterise a dynamic TB transmission model (Chapter \@ref(model-development)) and to fit the same model to data (Chapter \@ref(model-fitting)).

@@

### Demographic data

@@

```{r plot-overall-pop, fig.scap = "Overall population estimates derived using ONS and LFS demographic data.", fig.cap = "Overall population estimates derived using ONS and LFS demographic data. The ONS data is likely to be more reliable as the LFS data is derived using a weighted survey. After accounting for missing UK birth status both datasets provide comparable estimates of the population of England, with a clearly increasing trend over time. However the ONS data indicates a reduction in population from 2000 until 2001 that is not seen in the LFS data. The UK born population has also risen slowly from 2000 until 2015, although the biggest increase has been in the non-UK born population.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-overall-pop.png")
```

>>>>

Although the characteristics of tuberculosis (TB) in England have been reported elsewhere,[@PHE2017; @PHE2018] and key risk factors such as non-UK birth status have been identified,[@French2007] little attention has been given to the role of BCG vaccination. In particular, there is little information available regarding the demographics of vaccinated versus unvaccinated cases and the impact of BCG vaccination on TB outcomes in England has not been explored. There has also only been limited reporting of the age distribution, and trends over time, in incidence rates stratified by UK birth status.

In this chapter I explore the epidemiology of TB in England using routine data-sets with a particular focus on the impact of missing data, the mechanisms underlying that missing data, seasonal trends, the role of age, UK birth status and BCG status. I also estimate incidence rates, stratified by UK birth status and age, which I use to identify trends in TB incidence over time. Finally I report TB outcomes in England using case rates, again stratified by BCG status and UK birth status. These data are then used throughout this thesis to explore the impact of BCG vaccination on TB outcomes (Chapter \@ref(beneficial-bcg-out)), to estimate the direct impact of the 2005 change in BCG vaccination policy (Chapter \@ref(direct-eff)), to parameterise a dynamic TB transmission model (Chapter \@ref(model-development)) and to fit the same model to data (Chapter \@ref(model-fitting)).

@@

### Demographic data {#demographic-data}

@@

```{r plot-overall-pop, fig.scap = "Overall population estimates derived using ONS and LFS demographic data.", fig.cap = "Overall population estimates derived using ONS (Total) and LFS (Total (LFS)) demographic data. The ONS data is likely to be more reliable as the LFS data is derived using a weighted survey. After accounting for missing UK birth status both datasets provide comparable estimates of the population of England, with a clearly increasing trend over time. However the ONS data indicates a reduction in population from 2000 until 2001 that is not seen in the LFS data. The UK born and non-UK born populations are estimated using the LFS data.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-overall-pop.png")
```
